---
title: "Census Tract Event Rate and Community Deprivation"
author: "Cole Brokamp"
output: html_document
---

```{r echo=FALSE}
knitr::opts_chunk$set(prompt=F,comment=NA,warning=FALSE,message=FALSE,collapse=TRUE,cache=FALSE)
```

## Census API Key

Before you run this script, you will need to add your Census API key below.  


```{r, eval = TRUE}
tidycensus::census_api_key("PASTE YOUR KEY HERE", install = TRUE)
```

Setting `install =` to `TRUE` will store your API key in your `.Renviron` to so that it is available in future sessions.

## Importing Data

```{r}
library(tidyverse)
library(sf)

```

Read in case data coordinates file and convert to `sf` object.

```{r}
raw_data <- read_csv('./simulated_case_locations.csv')

d_events <- st_as_sf(raw_data, coords = c('X', 'Y')) %>% 
    st_set_crs(4326)
```

Quick check on event data.

```{r}
mapview::mapview(d_events)
```

Get the census tract geometries for Hamilton County.

```{r}
library(tigris)
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)
```

```{r}
hamilton_tracts <- tracts(state = 39, county = 061) %>% 
    select(GEOID)
```

Assign the corresponding census tract to each case location.  We will transform both data sets into the "Ohio South" projection system.

```{r}
d_events <- st_transform(d_events, 3735)
hamilton_tracts <- st_transform(hamilton_tracts, 3735)
d_events <- st_join(d_events, hamilton_tracts)
```

Summarize the number of events at the census tract level.

```{r}
d_tracts <- d_events %>% 
    group_by(GEOID) %>% 
    summarize(n_events = n()) %>% 
    st_set_geometry(NULL)
```

## Calculating the Event Rate

First, get the number of children per tract from the American Community Survey.

```{r}
d_pop <- tidycensus::get_acs(geography = 'tract',
                             variables = 'B09001_001E',
                             year = 2016,
                             state = 39, county = 61,
                             geometry = TRUE) %>%
    select(GEOID, n_children = estimate)
```

Merge in the number of events per tract and then calculate the event rate per 1,000 children in each tract.

```{r}
d <- left_join(d_pop, d_tracts, by = 'GEOID') %>% 
    mutate(event_rate = n_events / n_children * 1000)
```

Plot our event rate for each census tract.

```{r}
plot(d['event_rate'])
```

## Association with Deprivation Index

Merge in our deprivation index and look at the relationship with the event rate.

```{r}
d_dep <- 'https://github.com/cole-brokamp/dep_index/raw/master/ACS_deprivation_index_by_census_tracts.rds' %>%
  url() %>%
  gzcon() %>%
  readRDS() %>%
  as_tibble() %>%
  select(GEOID = census_tract_fips, dep_index)

d <- left_join(d, d_dep, by='GEOID')
```

```{r}
ggplot(d, aes(dep_index, event_rate)) +
    geom_point(alpha = 0.5)
```

Spearman's rank correlation between the tract-level deprivation index and event rate.

```{r}
cor.test(d$dep_index, d$event_rate, method = 'spearman')
```

## Creating Side By Side Publication Ready Maps

```{r}
library(tmap)
tmap_mode('plot')

d %>% 
    tm_shape(projection = 4326) +
    tm_polygons(c('event_rate', 'dep_index'))

```

Now let's pretty up our maps.

```{r}
d %>% 
    tm_shape(projection = 4326) +
        tm_polygons(c('event_rate', 'dep_index'),
                    alpha = 0.7,
                    style = 'cont',
                    palette = list('Blues', 'Reds'),
                    title = c('', '', '', ''),
                    showNA = FALSE,
                    lwd = 0.3,
                    border.col = "black", border.alpha = 1) +
    tm_compass(position=c('right', 'bottom')) +
    tm_scale_bar(position = c('right', 'bottom'), size=0.8) +
    tm_layout(title = c('Event Rate (per 1,000)',
                        'Deprivation Index'),
              title.position = c("left", "top"),
              title.size = 1.1,
              legend.position = c("left", "bottom"),
              legend.format = list(digits = 1),
              frame = FALSE,
              legend.text.size=0.8,
              inner.margins = c(0.11, 0.03, 0.12, 0.03), # bottom, left, top, right
              outer.margins = c(0, 0.05, 0, 0.05))
```

An interactive version of the maps.

```{r}
ttm()
tmap_last()
```

